---
title: "DS_NOTES_EDA"
output: github_document
---

```{r, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggridges)
library(patchwork)

```

```{r}
  weather_df = read.csv("C:/Users/Twilight/Desktop/Columbia/Fall Semester 2024/Data Science I/Datasets/weather_data.csv")
  weather_df <- weather_df %>% mutate(month = format(as.Date(date), "%m")) # creates a new column month extracted from date column
```

### Exploratory Data Analysis

* Exploratory Data Analysis is a loosely-defined process. Roughly, the stuff between loading data and formal analysis is exploratory.
* Grouping
  * Datasets often consists of groups(by design/implied/nested)
* Quantitative comparisons across groups are informative
  * Measures of Centrality, Variablity, Missingnes
* `group_by()` + `summarize()`
  * `group_by` makes grouping explicit and adds a layer to your data
  * `summarize()` allows you to compute one-number summaries
    * most useful in conjunction with `group_by()`


#### `group_by()`
```{r}
  weather_df %>%
    group_by(name, month)
```

### counting things
```{r}
  weather_df %>%
    group_by(month) %>%
    summarize(n_obs=n())

  weather_df %>%
    group_by(name, month) %>%
    summarize(count=n())
  
  # Using "count"
  weather_df %>%
    count(name, month, name="n_obs")
  
```











